### 2020-06-28 Report/aws_325ghz_study.tex
A first complete version!?

\documentclass[12pt]{article}

% Overall formatting
\sloppy
\let\cleardoublepage\clearpage % do not force chapters to start on odd pages

% Graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.png,.pdf,.jpg}
\graphicspath{{Figures/}}

\usepackage[a4paper, total={16cm,25cm}]{geometry}

% Tables
\newlength{\mytabskip}   
\setlength{\mytabskip}{-1.5ex}
\newcommand{\tabrule}{\rule[\mytabskip]{0ex}{0ex}}
\usepackage{tabularx}
%\usepackage{longtable}
\usepackage{mathtools}

% bibliography stuff 
\usepackage{natbib}
\bibliographystyle{kluwer}

\newcommand{\unit}[1]{\ensuremath{\mathrm{#1}}}
\newcommand{\degree}{\ensuremath{\mathrm{^\circ}}}

%\newcommand{\FIXME}[1]{{\sffamily \bfseries #1}}
\newcommand{\FIXME}[1]{{\sffamily {\bfseries \textcolor{red}{FIXME:} #1}}}

\hyphenation{EUMETSAT}
\hyphenation{Metop}

\begin{document}

% \thispagestyle{empty}
\pagenumbering{roman}

\noindent
\textbf{\Large Study to support the definition of Arctic Weather \vspace{1mm}\\
Satellite (AWS) high frequency channels} \vspace{8mm}\\
{\bf Patrick Eriksson, Inderpreet Kaur and Simon Pfreundschuh}\\
Department of Space, Earth and Environment\\
Chalmers University of Technology\\
SE-412\,96, Gothenburg, Sweden\vspace{10mm}

\section*{Summary}
%
The main objective of this study is to demonstrate that channels around
325\,GHz can provide a basis for cloud filtering/correction of data measured
around 183\,GHz. Previously, it was suggested that a channel at 229\,GHz to be
used for cloud filtering. The 229\,GHz channel forms a part of the MicroWave
Sounder (MWS) and will be used for cloud filtering of 183\,GHz inside that
mission. In this study, we investigate the possibility to use the 325\,GHz
channels instead. The cloud impact of high altitude cirrus clouds is
considerably higher in 325\,GHz and this makes it possible to estimate the weak
cloud impact at 183 GHz with relatively good accuracy.

On basis of the weighting functions of 183\,GHz, three specifications for
325\,GHz channels were suggested. A channel specific simple regression was
tested to estimate and correct the impact of hydrometeors in 183\,GHz channel.
The advantage of this approach is that for each 183\,GHz the cloud adjustment
can be quantified simply through a polynomial fit between the the cloud impact
in 183\,GHz channel of concern and brightness temperature differences between
183\,GHz and 325\,GHz. A similar regression scheme based on 229\,GHz was also
examined, but gave poor results. One drawback of the simple correction scheme
is its channel specific approach, which cannot benefit from the other 325\,GHz
channels available. In order to effectively use information from all 325\,GHz
channels, we suggested a machine learning based approach for cloud correction.
We used QRNN (Quantile Regression Neural Network) to to correct the cloud
impact in each 183\,GHz channel using measurements from all available channels
of 325\,GHz. All three options of 325\,GHz were evaluated, and the performance
was comparable. For some channels, interestingly, the spread of errors was
narrower than the noise of the target 183\,GHz channel. Another advantage of
using QRNN is that the cloud free values are predicted at different quantile
levels, which helps in quantifying the case-specific uncertainty estimate. A
similar QRNN model with 229\,GHz gave poor results. No critical dependency to
the 325\,GHz receiver noise temperature was observed.

This study shows that combining measurements from 325\,GHz can provide a better
cloud correction performance than 229\,GHz, and also with minimal rejection of
data. The clear-sky predictions have error distributions which are symmetric
with low bias and spread. Therefore, the NWP and standalone 1DVAR models that
utilise the ``clear-sky'' data can greatly benefit from it. Also, in data
assimilation systems, the case-specific uncertainty allows an effective
selection of weights, thus significantly improving the impact of measurements.
Another advantage of combining 183\,GHz and 325\,GHz for cloud correction is
that NWP systems, which are not yet prepared for 325 GHz, could still benefit
from the data early on.

The cloud correction analysis did not give a clear value on the relative merit
of the suggested channel options and also the measurements' degree of freedom
(DoF) was assessed based on the performed simulations. The DoF can be seen as a
measure of the effective number of channels. The DoF analysis gave further
support for selecting 325\,GHz in favour of 229\,GHz. The DoF of the 325\,GHz
options is higher or equal to the 229\,GHz one already in clear-sky conditions,
while for cloud conditions 325\,GHz consistently gives two additional degrees
of freedom. Some additional 325\,GHz channel versions were considered in the DoF
analysis.

With 325\,GHz as a better option confirmed, the next step is to establish
requirements for these channels. The study considered this question to some
extent. The cloud correction simulations gave no clear constraints on the
requirements, while the DoF analysis at least gave some indications. By
comparison to the NE$\Delta$D requirements for ICI and DoF values derived, it
is suggested to set threshold, target and breakthrough for the 325\,GHz
receiver noise temperature to 2400, 1800 and 1200\,K, respectively. Regarding
number, position and widths of channels, the differences between considered
options are relatively small and at this point it seems reasonable to let
technical considerations guide the basic selection between having three or four
channels. Later, when the technical constraints (such as the IF range
available) are better known an updated assessment should be performed to set
the final channel values.

\vspace{10mm}
\setcounter{tocdepth}{1} 
\tableofcontents

\newpage
\pagenumbering{arabic}
\setcounter{page}{1}

\section{Introduction}
%
\subsection{Background}
The Arctic Weather Satellite (AWS) is a small platform carrying a single
radiometer package. This instrument is an across-track scanning microwave
radiometer. The exact channel configuration of AWS is not yet set. The main
issue for this study is that AWS will have channels at 183\,GHz, but also one
or several channels a higher frequencies. A number of channels around 325\,GHz
was considered as the main option before AWS became an ESA/EUMETSAT mission,
but it has now also been suggested to have a single channel at 229\,GHz. This
suggestion is based on MWS (MicroWave Sounder), one of the instruments onboard
the next generation of Metop satellites. The 229\,GHz was added to MWS as an
help for performing cloud screening of 183\,GHz data
\citep{rekha2012potential}. For comparison, it can be mentioned that ICI (Ice
Cloud Imager), another new Metop sensor, will include 325\,GHz channels.

\subsection{Aim}
%
To support the selection between 229 and 325\,GHz channels, this report
investigates the possibility to use the latter range in forming 183\,GHz data of
``clear-sky'' character. That is, following \citet{rekha2012potential}, the
assumption, is that the 183\,GHz data will be used in an system not yet
capable of doing ``all -sky'' assimilation.

In \citet{rekha2012potential} the task was defined to screen out observations
having a cloud impact exceeding 4\,K at 183\,$\pm$\,7\,GHz. Their final suggestion
for screening does not involve data from any other channel than 229\,GHz, but
requires radiative transfer calculations (as a ``B-O test'' is made). Only 
screening of 183\,$\pm$\,7\,GHz was discussed.

In this report, the task is defined in a somewhat different manner. The aim is
not just to estimate the cloud impact, but also to investigate to which degree
the cloud impact can be removed, i.e.\ to make a cloud correction. Further, in
this report only schemes that can be based directly on AWS data will be
considered and versions involving ``B-O tests'' are left for possible future
studies.
%On the other hand, cloud correction of all AWS 183\,GHz channels are
%investigated. Another important difference is that our simulations aim to be a
%random sample of real situations, while the simulations of
%\citet{rekha2012potential} have a strong over-representation of cloudy
%cases.

\subsection{Comparison to \citet{rekha2012potential}}
\label{sec:rekha}
%
Results from \citet{rekha2012potential} will be used as reference. They did not
suggest specific settings of their algorithm. The main consideration of their
scheme is a threshold value. We assume below 6\,K for
this value, as it used most frequently as example in
\citet{rekha2012potential}. According to their Table~3, this results in that
2.4\% of the ``cloudy'' cases are miss-classified as ``clear''. We assume that
these cases have a cloud impact of 4 to 8\,K (i.e.\ the screening is 100\%
successful for cloud impact above 8\,K). For a 6\,K threshold 41\% of the
``clear'' cases are classified as ``cloudy'' and a removed in the screening
process.

The selected threshold value is towards the high side of the ones considered by 
\citet{rekha2012potential}. A lower value improves the removal of ``cloudy'' cases,
but the fraction of incorrectly removed ``clear'' cases increases quickly. For
example, a threshold of 3\,K decreases the miss-classification of ``cloudy''
cases to 0.6\%, but increases the miss-classification for ``clear'' cases to 77\%.
If 183\,$\pm$7\,GHz is classified as cloudy, we assume that also the data from
the other 183\,GHz channels are rejected.

\section{Assumptions on AWS characteristics}
%
The assumption listed below are primarily taken for the Statement of Work
(SoW), complemented with information provided by Anders Emrich at Omnisys
(private communication, April 14 2020). The later data represent what Omnisys
considers to be technically feasible inside the budget of the project.

\subsection{Channel specifications}
%
The specifications of channels around 166, 183 and 229\,GHz are assumed to
be fixed, with name, position and widths defined in Table~\ref{tab:fixed:chs}.
AWS channels 31 to 36 will be implemented as a single front-end, but are here
still considered as two ``bands'', denoted as 166 and 183\,GHz, to more easily
separate the channels in the text.

For 325\,GHz, it will be assumed that channels can be placed between 0.75 and
8.0\,GHz in terms of intermediate frequency (IF), and that up to 4 channels
inside this range can be implemented. There should be some gap between these
channels, above 100\,-\,200\,MHz, to facilitate the technical implementation.
For comparison, Table~\ref{tab:fixed:chs} includes also the position of 
325\,GHz channels of ICI. As the ICI channels extend to 11\,GHz in IF, they will not
be considered for AWS but will be applied for demonstration and reference
purposes.

For simplicity, the frequency variation of all channel responses is assumed to
be rectangular. This selection shall not be interpreted as any opinion on the
optimal shape of these responses; a significant deviation from a rectangular
shape should be of no concern as long as the actual shape is well
characterised.

\begin{table}[!b]
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-21  & \phantom{0}89.000 & 4000\\
    AWS-31  & 165.500 & 2800\\
    AWS-32  & 176.311 & 2000\\
    AWS-33  & 178.811 & 2000\\
    AWS-34  & 180.311 & 1000\\
    AWS-35  & 181.511 & 1000\\
    AWS-36  & 182.311 & \phantom{0}500\\
    AWS-4X  & 229.000 & 2000\\
    \hline
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    ICI-5  & 325.15$\pm$9.50 & 3000\\
    ICI-6  & 325.15$\pm$3.50 & 2400\\
    ICI-7  & 325.15$\pm$1.50 & 1600\\
    \hline
  \end{tabular}
  \end{minipage}  
  \caption{Fixed channel position and widths. The left table covers channels of
    AWS, with values taken from SoW. Channels AWS-31 to AWS-36 are at this
    moment mandatory for AWS, while AWS-4X is so far tentative. The right table
    lists the values of some channels of ICI \citep{eriksson:towar:20}.}
  \label{tab:fixed:chs}
\end{table}

\subsection{Noise}
%
The nominal receiver noise temperature, $T_r$ of each band is listed in
Table~\ref{tab:trec}. The noise of data for individual channels is calculated
as
\begin{equation}
  \sigma^i = c \frac{T^i_r+T^i_a}{\sqrt{\Delta\!f^i\Delta t}}
  \label{eq:noise}
\end{equation}
where $\sigma^i$ is the noise standard deviation of channel $i$, $c$ is a
scaling factor (see below), $T^i_r$ is the receiver noise temperature of the
channel, $T^i_a$ is the antenna temperature of the channel, $\Delta\!f^i$ is
the bandwidth (in terms of IF) of channel and $\Delta t$ is the integration
time.

The factor $c$ is set to 1.2 to incorporate the additional noise caused by the
calibration process. The bandwidth applied is the one found in columns 3 of
Table~\ref{tab:fixed:chs}. The integration time is set to match an effective
10\,km resolution at 183\,GHz. This time was estimated to 3\,ms.

\begin{table}[!bt]
  \centering
  \begin{tabular}[b]{c|ccccc}
    Band [GHz]   & 89 & 166 & 183 & 229 & 325 \\
    \hline
    $T_r$ [K]    & 390 & 650 & 650 & {\it 1000}  & 1200 \\
    \hline
  \end{tabular}
  \caption{Assumed receiver noise temperatures. The value for 229\,GHz is
    selected by us, remaining values are present assumptions at Omnisys.}
  \label{tab:trec}
\end{table}

\subsection{Various}
%
The AWS satellite is assumed to be at an altitude of 600\,km. Data were
generated for sensor viewing angles (the angle from nadir) from 0$^\circ$ to
45$^\circ$, in steps of 5$^\circ$.

The polarisation response of the bands is not yet determined and for simplicity
``quasi-horizontal'' (defined in same way as for e.g.\ MWS) is throughout
assumed. Antenna temperatures are obtained as:
\begin{equation}
  \label{eq:polrot}
  T_a = T_b^V\sin^2(\theta) + T_b^H\cos^2(\theta)
\end{equation}
where $T_b^V$/$T_b^H$ is the brightness temperature at vertical/horizontal
polarisation, respectively, and $\theta$ is the viewing angle (from nadir).

\section{Input data and software}

\subsection{Atmospheric scenarios}

\subsubsection{Fascod}
\label{sec:fascod}
%
The Fascod dataset \citep{anderson1986afgl} was used to represent different
climate zones in overview calculations. The dataset consists of profiles of
pressure, temperature and volume mixing ratio profiles of various atmospheric
gases. Five climatoligies were used: tropical (TRO), mid-latitude summer (MLS),
mid-latitude winter (MLW), sub-arctic summer (SAS), and sub-arctic winter
(SAW). No hydrometeors were included in simulations involving Fascod.

\subsubsection{Bulk profile database}
%
A database of atmospheric cases was created based on data from CloudSat and
ERA-Interim. The core idea of the approach is to use CloudSat reflectivities to
obtain as realistic vertical profiles of ice hydrometeors and rain as
possible. No external CloudSat retrievals are involved. Instead the CloudSat
reflectivities are mapped to properties for the passive simulations by assuming
a particle size distribution and a particle habit, together denoted as
the particle model. The process involves an inversion of the radar data, but as,
the same particle model is used for that inversion and to map ice water content
(IWC) and rain water content (RWC) to single scattering data when performing
the passive radiative transfer calculations, the impact of clouds on the
simulated AWS data is as realistic as possible. The thin clouds not detected by
CloudSat should not be of relevance for passive measurements frequencies
350\,GHz. Background data (temperature, humidity, 10\,m wind speed, \dots) were
taken from ERA-Interim, for the time and location of the CloudSat data. Liquid
water content (LWC) was taken from ERA-Interim, as CloudSat has no sensitivity
to the hydrometeor category.

This approach for generating atmospheric scenarios has been used by us since
\citet{rydberg:nonga:09}, and more recently in e.g.\
\citet{eriksson:towar:20} and \citet{barlakas:three:20}. It is also used in
\citet{ekelund:using:20}, where the approach is described more in detail. In
the cited articles, the sensor's footprint has been considered in varying
degree, as ``beam filling'' must be considered at high impact of hydrometeors.
As clear-sky and weak cloud interference are in focus in this study, and to
save calculation time, instead CloudSat data were averaged over 10\,km to form
the atmospheric cases and standard 1D radiative transfer calculations are
applied on each case (instead of applying an independent beam approximation
inside a 2D or 3D atmosphere as done in cited works).

\subsection{Radiative transfer simulations}
%
\subsubsection{ARTS}
%
All radiative transfer calculations are made by the Atmospheric Radiative
Transfer System (ARTS, \citet{eriksson:arts2:11,buehler:artst:18}), version
arts-2.3.1284. Models and input used for absorption and hydrometeor properties
are described below. Ocean/water and land emissivities are taken from
TESSEM \citep{prigent2017sea} and TELSEM \citep{aires2011tool}, respectively.

Two calculations were performed for each atmospheric cases. One ''clear-sky''
where all hydrometeors contents were set to zero, and one ``all-sky'' where IWC
and RWC derived from CloudSat reflectivities and LWC from ERA-Interim were
included. Both sets of calculations were made by ARTS's interface to the
RT4 solver \citep{evans1995microwavec}. This ``scattering solver'' was used
also for clear-sky, to avoid a possible bias between clear-sky and all-sky for
insignificant hydrometeor contents. RT4 handles the first two elements of the
Stokes vector, that were converted to brightness temperatures for H- and
V-polarisation.

\subsubsection{Absorption models}
%
The absorption of nitrogen is modelled following \citet{pwr:93}. Nitrogen has a
significant impact at least around 325\,GHz at dry conditions. Also oxygen
absorption follows \citet{pwr:93}. Oxygen is in this study mainly of concern
for the 94\,GHz CloudSat simulations. LWC was treated to be
purely absorbing, based on \citet{ellison2007permittivity}.

For water vapour the present settings in RTTOV are applied. That is, largely
MPM89 \citep{liebe:89} is followed, but some parameters for the 22 and 183\,GHz
transitions are replaced \citep{saunders2018update,turner2019amsutran}. A
smaller comparison to RTTOV has been made, that indicated an agreement better
than 0.1\,K in general. Deviations between 0.1 and 0.2\,K were found for some
channels at incidence angles above 40$^\circ$.

\subsubsection{Hydrometeor properties}
%
As mentioned above, LWC is taken from ERA-Interim and is assumed to be totally
absorbing. LWC is allowed to exist at temperatures below 0$^\circ$C, i.e.\
super-cooled liquid cloud droplets are present in the simulations. In the
mapping of CloudSat reflectivities to RWC and IWC a total separation between
liquid and ice phase is assumed. All scattering hydrometeors at temperatures
above 0$^\circ$C are assumed to be rain, and all below 0$^\circ$C are assumed
to be ice hydrometeors.

For RWC the particle size distribution of \citet{abel2012improved} is applied.
The PSD of IWC follows the basic formulation applied in DARDAR, using latest
parameter values (i.e.\ $\alpha$ and $\beta$) as given by
\citet{amt-12-2819-2019}. This PSD can be considered as a ``two moment''
scheme, but is here applied in a one moment manner by setting $N_0^*$ (as a
function of temperature) following Table~5 of \citet{delanoe2014normalized},
and letting the radar reflectivity set the remaining moment.

Single scattering data are taken from \citet{eriksson:agene:18}. For ice
hydrometeors, three habits are applied: Perpendicular 3-bullet rosette, Large
plate aggregate and Large column aggregate. In the last two cases, the
aggregates are complemented with single crystal data to also cover smaller
sizes. These data describe particles assumed to have a totally random
orientation. To apply oriented particles is much more computationally costly
and could not be accommodated inside the study.

\subsection{Dataset formed}
%
Using the simulation setup described in previous section, we simulated 125460 cases based on randomly selected CloudSat orbits during August 2015. The input data
were restricted between $60^{\degree}$S to $60^{\degree}$N, and surface is below 500\,m. Both clear-sky and all-sky scenarios were simulated, and no differentiation was made between the observations over water and land. 24 monochromatic frequencies were used so that each channel could be represented with sufficient accuracy. The simulations were generated for sensor viewing angles from $0^\circ$ to $45^\circ$, but all the results described in this report are based on nadir viewing angle.

Since in this study, the cases with cloud impact are extremely important, we used ``clear-sky thinning'' to improve the representation of the cloudy cases in the dataset. Thinned dataset is a subset of the full dataset with 50\% of the clear-sky cases removed (cloud impact $<$ 0.2\,K) The size of the thinned dataset is different for each channel.

\subsection{Validation of simulations}
\label{sec:validation}
%
\begin{figure}[p]
	\centering
	\includegraphics[height=60mm]{ATMS_C16_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C17_distribution}
	\includegraphics[height=60mm]{ATMS_C18_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C19_distribution}
	\includegraphics[height=60mm]{ATMS_C20_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C21_distribution}
	\includegraphics[height=60mm]{ATMS_C22_distribution}
	\caption{Probability distribution functions (PDFs) of simulated and
      observed ATMS channels brightness temperatures for channels 16 - 22 and
      nadir viewing angle. The data covers latitude range $60^{\degree}$S to
      $60^{\degree}$N. Both land and ocean measurements are considered and only
      daytime ATMS data are included. See further the text.}
	\label{fig:pdf:c16-22}
\end{figure}

A statistical comparison of the simulations to actual ATMS data, for channels
16-22 at nadir (zenith angle = $180^{\degree}$), is shown in
Fig.~\ref{fig:pdf:c16-22}. For the comparison, the daytime ATMS observations
were selected as CloudSat operates only during the day part of the orbit since
2011. For the channels around 183\,GHz, the ATMS characteristics are not fully
simulated. The ATMS channel widths are considered, but as AWS is only covering
the lower wing of the 183\,GHz transition the upper sideband of ATMS is not
included.

The overall agreement is good for all channels. For 183 GHz channels (Channels
18-22), the simulations and observations have a good agreement for the part of
the PDFs matching clear-sky conditions (the main peak of the distribution),
albeit the simulations have a more sharp peak. This could either be caused by a
different geographical sampling of ATMS and CloudSat or to shortcomings in the
ERA-Interim input data. The fraction of cloudy observations is lower in ATMS in
comparison to simulations, for brightness temperatures below about 240\.K. This
is due to the fact that the antenna pattern is not considered in the
simulations.

In the intermediate region, the simulations have a lower occurrence rate. The
impact of hydrometeors should be somewhat higher in the upper sideband of ATMS,
which is not reflected in the simulations, but this is not sufficient to
explain the deviation. As this region is well covered in the simulations
reported in Fig.~7 of \citet{eriksson:towar:20}, the deviation of concern is
presumably also a consequence of the neglected antenna pattern. In this study,
this transitional region is of high interest. For this reason, the PDF for a
50\% random rejection of cases with an insignificant cloud impact is also
shown. This PDF gives a fair match in the intermediate region (but gives an
even higher over-representation at more strong cloud impacts). This smaller
dataset will be considered in parallel below, to ensure that no fraction of the
``cloudy'' cases is not underrepresented in the results.

\section{Overview of the 183, 229 and 325\,GHz bands}
\label{sec:overview}
% 
The simulations in this section are based on five Fascod climate scenarios
(Sec.~\ref{sec:fascod}), and performed with a fixed sensor viewing angle of
25$^\circ$ (deviation from nadir). No hydrometeors are included. For simpler
interpretation of the results, the reflectivity is set to have no frequency
variation, i.e.\ the same value is applied for all bands. The reflectivity is
varied between 0 and 0.5 (these fixed values are only applied in this section,
the bulk simulations are using TESSEM and TELSEM for setting the surface
properties). In the figures covering transmissivities and Jacobians, the
specifications of ICI is applied for 325\,GHz channels
(Table~\ref{tab:fixed:chs}).

\subsection{Clear-sky brightness temperatures}
%
Figure~\ref{fig:tb:r000} shows brightness temperatures over the three bands, for
a reflectivity of 0 (i.e.\ a blackbody surface). The figure takes into account
that the 183\,GHz channels of AWS are of single-sideband character, while for
229 and 325\,GHz averages of upper and lower sidebands are shown, plotted as
a function from the centre frequency on the lower side.

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r000}
  \caption{Solid lines show brightness temperatures for the lower wing of the
    183.31\,GHz transition. Dashed lines represent double-sideband measurements
    around 325.15\,GHz, with the mean of upper and lower band shown on the low
    side. Dash-dotted lines represent double-sideband measurements around
    229.00\,GHz in the same manner, but only over the bandwidth of the
    potential channel. For all three bands, spectra for all five
    Fascod scenarios are included.}
  \label{fig:tb:r000}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r050}
  \caption{Same as figure above, but with a surface reflectivity of 0.5.}
  \label{fig:tb:r050}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tr_mls}
  \caption{Transmissivity to space as a function of altitude, for the
    mid-latitude summer scenario. Solid lines represent 183\,GHz channels,
    dashed lines 325\,GHz channels and dashed-dotted the 229\,GHz channel
    (Table~\ref{tab:fixed:chs}). These transmissivities are independent of
    surface reflectivity.}
  \label{fig:tr:mls}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_wf_mls_025}
  \caption{Jacobians for relative humidity. Line symbols as in Fig.~\ref{fig:tr:mls}.
    Surface reflectivity is set to 0.25.}
  \label{fig:wf:mls:025}
\end{figure}

For 183 and 325\,GHz the brightness temperatures increase when moving away from
the centre frequency of the transition. The ``kinks'' in the spectra correspond
to ozone transitions, that are both more frequent and stronger in the 325\,GHz
range than in the 183\,GHz one. See Fig.~1 of \citet{eriksson:towar:20} for how
the ozone transitions are distributed between the lower and upper 325\,GHz
sideband. The spectra of 183 and 325\,GHz deviate little close to the transition
frequency, showing that the two frequencies are of similar strength, though the
325\,GHz one is slightly weaker. The higher deviations further out in the wings
are due to different contribution of far wing and continuum absorption, that is
higher at 325\,GHz. The 229\,GHz brightness temperatures are higher than any
value around 183 and 325\,GHz, as expected 229\,GHz being a window channel.

In Fig.~\ref{fig:tb:r050} the surface reflectivity is changed to 0.5, which
should roughly correspond to the highest reflectivity encountered at these
frequencies. The differences to Fig.~\ref{fig:tb:r000}  are small for 183 and
325\,GHz for the tropical and the two summer scenarios. On the other hand,
there is a clear impact in the wing part of 183\,GHz for the two winter
scenarios. The same is true for 325\,GHz, but the influence of the surface
reflectivity is considerably lower.  Beside for the tropical case, for
229\,GHz there is a pronounced dependency of the surface reflectivity, e.g.\
about 35\,K for mid-latitude winter.

\subsection{Clear-sky transmissivities to space}
%
Assuming that a cloud layer can be moved in altitude, its impact on measured
brightness temperatures will roughly follow the transmissivity to space for the
cloud altitude. In short, where the transmissivity is zero the cloud will have
no impact, while where the transmissivity is unity the influence will be close
to constant with altitude (especially when the single scattering albedo is
high). This relationship just refers to the relative impact for a
given frequency, the absolute impact (at transmissivity 1) increases in general
with frequency.

Figure~\ref{fig:tr:mls} displays examples on transmissivities to space. The
transmissivities refers to the up-welling radiation. The contribution from
reflected down-welling is neglected and the considered transmissivity does not
depend on surface reflectivity.

The altitude variation of transmissivity exhibits same basic shape between all
the 183 and 325\,GHz channels, but the altitude of optical thickness of 1
(transmissivity of $e^{-1}\approx0.37$) varies. The curves of the edge 183\,GHz
channels (AWS-32 and 36) brackets the ones corresponding to the ICI 325\,GHZ
channels. The later channels show a higher deviation from 1 at 150\,hPa, due to
some (weak) ozone attenuation in the stratosphere. The 229\,GHz has the highest
transmissivities.

\subsection{Clear-sky relative humidity weighting functions}
%
The sensitivity to humidity is most easily compared by the channels' weighting
functions (i.e.\ the Jacobian), and examples are found in
Fig.~\ref{fig:wf:mls:025}. The shape of the weighting functions depends on the
unity selected for representing water vapour. As the absolute amount of water
vapour varies with orders of magnitude from the surface to the tropopause, it
is difficult to interpret weighting functions for absolute humidity units and
it is more suitable to use a relative one. In Fig.~\ref{fig:wf:mls:025} the
weighting functions correspond to an increase of relative humidity (RH) with one
percent unit over 1\,km. RH with respect liquid/ice is applied above/below
0$^\circ$C.

The basic shape of the Jacobian is similar for all 183 and 325\,GHz channels,
but the peak altitude varies. Again the edge 183\,GHz channels brackets the
325\,GHz ICI ones, as for transmissivity. These Jacobians are negative, meaning
that a change towards higher humidity results in decreased brightness
temperatures. The 229\,GHz Jacobian is positive close to the surface, as this
channel has sensitivity all the way down to the surface and more water vapour
just above a reflecting surface results in higher brightness temperatures. The
switch from negative to positive Jacobian depends on atmospheric scenario and
surface reflectivity. For dry conditions also the outer 183 and 325\,GHz
channels can have a positive part, as long as the reflectivity is not close to
0.

\section{Preliminary options for 325\,GHz channels}

\label{sec:channel:configurations}

\subsection{Impact of upper end of IF band}
%
Figure~\ref{fig:wfuns:325ul} assumes that the channel placed at the high end of
the IF band is 3\,GHz wide (as for ICI, Table~\ref{tab:fixed:chs}), and shows
the RH weighting function if the outer channel ends at 11, 10, 9 or 8\,GHz. The
weighting functions for each sideband are here kept separated.

As expected, when the channel is moved away the centre frequency, the weighting
function moves downwards. However, moving the channel from 5\,-\,8\,GHz to
8\,-\,11\,GHz (in IF) results only a small shift, about 400 and 200\,m for the
lower and upper side, respectively. In terms of the resulting double sideband
channel the the shift would be 300\,m. Accordingly, there is little gain by
enforcing the IF range to go above 8\,GHz. As a rule of thumb, a higher ratio
between highest and lowest IF is more costly to implement. Albeit small, a
negative consequence of a 8\,-\,11\,GHz channel, compared to 5\,-\,8\,GHz, is
some asymmetry in the altitude response between lower and upper sideband. This
asymmetry is caused by higher continuum and far wing absorption on the upper
side.

In summary, the benefit of extending the IF band above 8\,GHz is small and
placing channels above 8\,GHz seems unnecessary considering the limited budget
and development time of AWS.

\begin{figure}[!p]
  \centering
  \includegraphics[height=61mm]{fascod_wf_325l_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=61mm]{fascod_wf_325u_tro}
  \caption{RH weighting functions for the lower (left) and upper (right) side
    of the 325.15\,GHz transition. All weighting functions are for a 3\,GHz
    bandwidth. The stated frequencies are the centre frequency of each. In
    terms of IF range, the channels cover 8-11, 7-10, 6-9 and 5-8\,GHz (from
    top-to-bottom in the order given in the legends). }
  \label{fig:wfuns:325ul}
\end{figure}
\begin{figure}[!p]
  \centering
  \includegraphics[height=61mm]{fascod_3chopt_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=61mm]{fascod_3chopt_mlw}
  \caption{RH weighting functions of the three channel option in
    Table~\ref{tab:chs:prel3}, for the Fascod tropical (left) and mid-latitude
    winter (right) scenarios.}
  \label{fig:3ch:prel}
\end{figure}
\begin{figure}[!p]
  \centering
  \includegraphics[height=61mm]{fascod_4chopt_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=61mm]{fascod_4chopt_mlw}
  \caption{RH weighting functions of the four channel option in
    Table~\ref{tab:chs:prel4}, for the Fascod tropical (left) and mid-latitude
    winter (right) scenarios.}
  \label{fig:4ch:prel}
\end{figure}

\subsection{A three channel option}
%
The ICI channels give a rough RH weighting functions matching with AWS-32,
AWS-33 and AWS-34 (Fig.~\ref{fig:wf:mls:025}), and to find a tentative three
channel option an even better matching was used as constrain to set the
specifications. The resulting values are found in Table~\ref{tab:chs:prel3} and
example weighting functions are displayed in Fig.~\ref {fig:3ch:prel}. This
option has a IF range of 1.0 to 8.0\,GHz, and leaves a 100\,MHz space between
AWS-42 and AWS-43, and 200\,MHz space between AWS-41 and AWS-42. That is, the
suggestion uses the complete IF bandwidth as far as possible, that is
technically possible.

\subsection{A four channel option}
%
The tentative four channel option in Table~\ref{tab:chs:prel4} gives a broader
altitude coverage, towards higher altitudes, and offers weighting functions
roughly equally spaced in altitude (Fig.~\ref {fig:4ch:prel}). This option has
a IF range of 0.7 to 8.0\,GHz, and leaves 200\,MHz space between all the
325\,GHz channels. Also this suggestion uses the complete IF bandwidth.

\begin{table}[!p]
  \centering  
  \begin{tabular}[c]{c|c|c|ccc}
    Channel & Frequency   & Bandwidth & $T_r=1200$\,K
    & $T_r=1800$\,K & $T_r=2400$\,K\\
    name    & [GHz] &  [MHz] & [K]  & [K] & [K]\\
    \hline
    AWS-41  & 325.15$\pm$6.50 & 3000 & 0.58 & 0.82 & 1.06 \\
    AWS-42  & 325.15$\pm$3.55 & 2500 & 0.64 & 0.90 & 1.16 \\
    AWS-43  & 325.15$\pm$1.60 & 1200 & 0.92 & 1.30 & 1.68 \\
    \hline
  \end{tabular}
  \caption{Tentative suggestion for a three 325\,GHz channels option (three-a). The
    values in the three last columns is the 1$\sigma$ noise level according to
    Eq.~\ref{eq:noise}, for $T_a=300$\,K and stated $T_r$ }
  \label{tab:chs:prel3}
\end{table}

\begin{table}[!p]
  \centering  
  \begin{tabular}[c]{c|c|c|ccc}
    Channel & Frequency   & Bandwidth & $T_r=1200$\,K
    & $T_r=1800$\,K & $T_r=2400$\,K\\
    name    & [GHz] &  [MHz] & [K]  & [K] & [K]\\
    \hline
    AWS-41  & 325.15$\pm$6.60 & 2800 & 0.60 & 0.85 & 1.10 \\
    AWS-42  & 325.15$\pm$4.10 & 1800 & 0.75 & 1.06 & 1.37 \\
    AWS-43  & 325.15$\pm$2.40 & 1200 & 0.92 & 1.30 & 1.68 \\
    AWS-44  & 325.15$\pm$1.20 & \phantom{0}800 & 1.12 & 1.59 & 2.06\\
    \hline
  \end{tabular}
  \caption{Tentative suggestion for a four 325\,GHz channels option. Else as
    Table~\ref{tab:chs:prel4}.}
  \label{tab:chs:prel4}
\end{table}

\begin{table}[!p]
  \centering  
  \begin{tabular}[c]{c|c|c|ccc}
    Channel & Frequency   & Bandwidth & $T_r=1200$\,K
    & $T_r=1800$\,K & $T_r=2400$\,K\\
    name    & [GHz] &  [MHz] & [K]  & [K] & [K]\\
    \hline
    AWS-41  & 325.15$\pm$6.50 & 3000  & 0.58 & 0.82 & 1.06\\
    AWS-42  & 325.15$\pm$3.30 & 3000  & 0.58 & 0.82 & 1.06\\
    AWS-43  & 325.15$\pm$1.20 & \phantom{0}800 & 1.12 & 1.59 & 2.06\\
    \hline
  \end{tabular}
  \caption{The second three channel suggestion (three-b).}
  \label{tab:chs:3b}
\end{table}

\begin{figure}[!p]
  \centering
  \includegraphics[height=61mm]{fascod_3bchopt_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=61mm]{fascod_3bchopt_mlw}
  \caption{RH weighting functions of the second three channel option
    (Table~\ref{tab:chs:3b}), for the Fascod tropical (left) and mid-latitude
    winter (right) scenarios.}
  \label{fig:3bch:prel}
\end{figure}

\subsection{A second three channel option}
%
Relatively late in the study Omnisys informed us about that an IF of 1.7\,GHz
should be avoided, as this frequency likely will be used for downlink of data.
The four channel option needed just a marginal change to adopt to this
constrain, while 1.7\,GHz is more or less at the centre of the three channel
option. It was then decided to formulate second three channel option avoiding 
1.7\,GHz, and the suggestion is found in Table~\ref{tab:chs:3b}.

AWS-41 is identical between the two three channel options. In this second
option, AWS-43 is set to have the same limits as AWS-44 in the four channel
option. AWS-42 was then allowed to fill the frequency range between the two
other channels, but keeping a gap of 200\,MHz on each side.

The resulting weighting functions are exemplified in Fig.~\ref{fig:3bch:prel}.
As expected, there is now a somewhat poorer match to 183\,GHz, most importantly
the weighting function of AWS-43 ends up between the ones of AWS-35 and 36. The
weighting function of AWS-42 is slightly moved upwards compared to the first
three channel option.

\section{Cloud correction by a simple scheme}
\label{sec:simple}
%
In this section, the (first) 325\,GHz three channel option is evaluated for detection
and correction of cloud interference in 183\,GHz channels. Combinations of
brightness temperature (TB) differences between 183\,GHz and 325\,GHz are used
as a measure of the cloud impact. A regression based approach is used to
estimate the adjustment needed to predict the clear-sky values.

\subsection{Formulation of correction scheme}
\label{sec:correction:scheme}
%
Let C1 and C2 represent one of the channels each from 183\,GHz and 325\,GHz,
respectively, and Tb1 and Tb2 represent the corresponding TB. The correction
scheme formulates a relationship between the cloud impact in C1 to the TB
differences between C1 and C2:
\begin{equation}
Tb1_{as}-Tb1_{cs} = f(x_m)
\label{eq:TB:diff}
\end{equation}
where, $x_m = Tb2_{as} - Tb1_{as}$, and the subscripts ``as'' and ``cs'' denote the all-sky and clear-sky cases respectively.

The combinations of C1 and C2 are selected by finding best match between
weighting functions. Five 183\,GHz and 325\,GHz combinations are possible, for
example one such paired combination is AWS-34 and AWS-42.

\begin{figure}[!b]
	\centering
	\includegraphics[height=85mm]{fit_AWS-34_AWS-42}\hspace{5mm}% 
	\caption{Brightness temperature difference of AWS-34 from AWS-42 plotted
      against cloud impact in AWS-34. Red and green colour represent the clear
      and cloudy cases respectively in the simulation dataset. The black curve
      represents the line of best fit (of green points) as a third order
      polynomial. The blue vertical line represents the correction threshold,
      $x_{cs}$, for $dtb_{cs} =$ 0.2\,K. }
	\label{fig:fit:c34-42}
\end{figure}
%

As an example, the relationship between the TB differences for C1 and C2 and
the cloud impact in C1 is shown in Fig.~\ref{fig:fit:c34-42}. For this
particular case, C1 is AWS-34 and C2 is AWS-42, and around 125\,000 simulations
are used. The two variables have a reasonably strong correlation for small TB
differences, but as the impact of cloud increases, the correlations gets
poorer. The data can be fitted by linear or non-linear models. The fitted model
can be used to obtain the adjustment required to compensate the brightness
temperature for cloudy cases. To avoid fitting the ``too cloudy'' cases which
do not show any reasonable correlation, cases with $Tb2-Tb1< -40$\,K are
excluded in the fit. Also, cases with a cloud impact below 0.2\,K (red colour
dots) can be considered as clear-sky, are not included in the model.

It is also important to have a manner for making a clear/cloudy classification
that can be derived from actual measurement data. This is useful for making
filtering in line with \citet{rekha2012potential}, but is also needed in the
correction scheme. The latter is the case as the polynomial fit derived gives a
non-zero correction for clear-sky cases. In Fig.~\ref{fig:fit:c34-42} the
difference between clear-sky values of channel C1 and C2 (red dots) have a mean
close to zero (as the weighting functions are well matched), but this is not
the case for all channel combinations and some general rule for finding a
threshold value (with respect to $x_m$) must be formulated. The polynomial fit
will be used also for this part.

We will simply treat cases that matches a correction below a certain value as
``clear-sky''. If the correction limit is denoted as $dtb_{cs}$, the threshold
value, $x_{cs}$, is obtained by inverting the polynomial fit:
\begin{equation}
x_{cs} = f_{cs}^{-1}(dtb_{cs}).
\label{eq:dtb}
\end{equation}
That is, if $x_m>x_{cs}$ a measurement is classified as clear-sky. If
$x_m\leq x_{cs}$, the measurement is rejected if a pure filtering is made, or
it is adjusted if the correction scheme is applied. We will largely set
$dtb_{cs}$ based on the noise level of the 183\,GHz channel involved, to be
1$\sigma$ or 2$\sigma$ where $\sigma$ refers to the magnitude of measurement
noise (Eq.~\ref{eq:noise}, often called NE$\Delta$T). We will use $1\sigma$ for
the development of the scheme, however, the sensitivity of the scheme to
a 1$\sigma$ or 2$\sigma$ threshold is also provided later in the text.

\subsection{Application of correction scheme}
%
In this section, we test the applicability of the correction scheme in detail.
To analyse the performance of the approach, we apply it to correct the
measurements from C1 using data from C2. Without the availability of actual
data, the simulation dataset but with noise added is used mimic the
measurements. The noise for individual channels is calculated according to
Equation~\ref{eq:noise}. As an example, the results from the channel
combination AWS-34 and AWS-42 are shown. The error in the predicted
measurements is assessed as the deviation to corresponding noise-free clear-sky
value. The average bias, standard deviation and the skewness of the
deviations are calculated. Skewness gives a measure of asymmetry of the
probability distribution of deviations around its mean.

To test the efficacy of the correction scheme, three different models: linear,
quadratic and cubic are evaluated to fit the TB differences between C1 and C2.
For all three models, cases with $Tb2-Tb1 < $ -15\,K are assumed to be ``too
cloudy'' and are not corrected. Also, cases classified as clear according to
the $1\sigma$ threshold are not corrected. The clear-sky adjustment for all
eligible measurements is calculated from all three models, and the three
predicted datasets are compared.

The occurrence frequency of the deviations from three models is shown in
Fig.~\ref{fig:correction:c34-42:fit} and their corresponding statistics are shown in Table~\ref{tab:correction:stats:34:42:allfits}. The occurrence frequency of the
measurements without any filtering or correction is also plotted as reference
(orange colour). If all measurements are devoid of cloud impact, the
deviations should follow the Gaussian distribution of noise. In such a case, the skewness of the deviations is close to zero.  However, in real conditions, the distribution has a large negative skewness due to the impact of clouds. After correction, all the three models are successful in reducing the cloud impact in the
measurements. The linear model has the lowest reduction in bias (from  -0.75\,K to -0.03\,K) but the highest values for skewness and standard deviation, in comparison to the other two models.  With the linear model, the error distribution of the corrected dataset has a relatively high positive skewness, which is undesirable. The positive skewness can be due to over-correction of cloudy cases, or erroneous correction of clear cases. The cubic  and quadratic models have almost similar statistics for bias and standard deviation, but the skewness in the cubic model is lower. Overall, the cubic model can be said to have the best performance, thus for all the results shown further, a cubic model is used.

In Sec.~\ref{sec:validation} the validation of simulations against ATMS data
(Fig.\ref{fig:pdf:c16-22}) showed that the cloudy cases in the simulations were
underrepresented in comparison to the observations. In order to have a higher
representation of cloudy cases, we thinned the measurement dataset by removing
50\% of the clear-sky cases. The distribution of thinned dataset is more closer
to real conditions in the region matching weak cloud impacts and all the
analyses presented further in this section are based on the thinned dataset.

Figure~\ref{fig:correction:c34-42:thinned} shows the deviation of the uncorrected and
corrected measurements to their noise-free clear-sky values for the thinned dataset. The results from \cite{rekha2012potential} are also shown as reference (see Sec.~\ref{sec:rekha}
for our assumptions). In order to allow a more direct comparison of our
approach with \cite{rekha2012potential}, we also include the
dataset, ``Filtered'', where all cloudy cases are rejected. Apart from the fit
based cloudy classification, it is also assumed that cases with $Tb2-Tb1 < $
-15\,K are ``too-cloudy'' and are rejected.  To analyse the sensitivity of our scheme to noise threshold, as discussed in the previous section, results from two different values of $dtb_{cs}$ are shown. For both values of $dtb_{cs}$ the corresponding clear-sky cut-off
($x_{cs}$) is calculated using Eq~\ref{eq:dtb}. Various metrics corresponding to
the Figure~\ref{fig:correction:c34-42:thinned} are shown in
Table~\ref{tab:correction:stats:34:42:thinned}.

\begin{figure}[!tb]
	\centering
	\includegraphics[height=100mm]{PDF_corrected_AWS-34_AWS-42_allfits}
	\caption{Deviation of AWS-34 measurements to noise free clear-sky
      simulations before and after correction, corrected using a linear,
      quadratic and cubic model. The distribution of noise alone is also shown
      as reference. The measurements have been corrected using brightness
      temperatures from AWS-42. The clear/cloudy classification is made with $dtb_{cs} = 1\sigma$ and also cases with $Tb2 - Tb1 > -15$\,K are rejected.}
	\label{fig:correction:c34-42:fit}
\end{figure}
%
\begin{table}[!tb]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness  & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		noise                     &   0.00 &  0.63 &               0.00 &                - \\
		uncorrected-all           &  -0.75 &  4.62 &             -11.78 &                - \\
		corrected-linear model    &  -0.03 &  0.75 &               0.48 &                2.84 \\
		corrected-quadratic model &  -0.06 &  0.71 &              -0.34 &                2.84 \\
		corrected-cubic model     &  -0.05 &  0.72 &              -0.23 &                2.84 \\
		\hline
	\end{tabular}
	\caption{Bias, standard deviation (std), and measure of skewness of error
      distributions shown in Figure~\ref{fig:correction:c34-42:fit}. The
      fraction rejected describes the percentage of cases which are removed.
      Results from three different fit models are shown: linear, quadratic and
      cubic.}
	\label{tab:correction:stats:34:42:allfits}
\end{table}
%
\begin{figure}[!tb]
  \centering \includegraphics[height=100mm]{PDF_corrected_AWS-34_AWS-42_thinned}
	\caption{Deviation of AWS-34 measurements to noise free clear-sky
      simulations before and after correction. The measurements have been
      corrected using a cubic model. The different datasets are: noise
      (simulation dataset), uncorrected (thinned measurements), filtered $1\sigma$ (cases classified as cloudy according to $dtb_{cs}$ = 1$\sigma$ are rejected), corrected $1\sigma$ (cases classified as cloudy according to $dtb_{cs}$ = 1$\sigma$ are corrected), corrected
      $2\sigma$ (cases classified as cloudy according to $dtb_{cs}$ = 2$\sigma$ are corrected), Sreerekha et al (cloudy cases are filtered according to \citet{rekha2012potential}).}
	\label{fig:correction:c34-42:thinned}
\end{figure}
%
\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness  & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
	noise             		&   0.00 &  0.63 &              -0.01 &                - \\
	uncorrected     		&  -1.27 &  6.05 &              -8.97 &                - \\
	filtered($1\sigma$)  	&  -0.13 &  0.68 &              -0.25 &               17.54 \\
	corrected($1\sigma$) 	&  -0.06 &  0.76 &              -0.46 &                4.99 \\
	corrected($2\sigma$) 	&  -0.11 &  0.75 &              -0.69 &                4.99 \\
	sreerekha et al         &  -0.16 &  0.71 &              -0.46 &               46.37 \\			
		\hline
\end{tabular}
\caption{Bias, standard deviation (std), and measure of skewness of the error
  distributions of datasets shown in
  Figure~\ref{fig:correction:c34-42:thinned}.}
\label{tab:correction:stats:34:42:thinned}
\end{table}

The distributions of filtered data are non-Gaussian. The right side of the
distribution follows measurement noise while the left side is skewed. The
skewness in the distribution is due to cloud impacts which pass through the
clear-sky threshold. The skewness in ``Sreerekha et al'' is higher than our approach because their method only aims at removing cloud impacts above 4\,K. Their approach also leads to a high rejection fraction (46.37\%). The filtering
method introduced here leads to a considerable lower rejection of data
(17.54\%).

If the correction part for $1\sigma$ is added, the rejection is decreased even further (4.99\%)
and the bias is also improved (from -1.27 to -0.06\,K). On the other hand, the
correction gives some increase in standard deviation and skewness, compared to
filtering alone. A correction using $2\sigma$ gives poorer performance, in particular the reduction in bias is smaller and the skewness is increased.

%
\subsection{Application of correction scheme using AWS-4X}
%

As a reference with respect to having a channel at 229\,GHz, we also apply the
correction scheme described above to correct AWS-32 using AWS-4X (following
\citet{rekha2012potential}). In this case, C1 and C2 are AWS-32 and AWS-4X. The
TB correlations between AWS-32 and AWS-4X are poor and show a large spread
particularly for large cloud impacts (not shown). We fit the data to a polynomial model as described in Sec.~\ref{sec:correction:scheme} but exclude cases with $Tb2-Tb1 <$ -20\,K in the fit.

Fig.~\ref{fig:correction:c32-4X:thinned} shows the deviations of the
measurements to clear-sky values before and after correction, and the
corresponding metrics are provided in Table~\ref{tab:correction:stats:32:4X}.
With clear/cloudy classification for $dtb_{cs} = 1\sigma$, around 85\% of the
total measurements are classified as cloudy and are rejected in pure-filtering.
Rejecting all the cloudy cases reduces the average bias in the measurements
from -2.25\,K to -0.21\,K, while with the approach from
\citet{rekha2012potential} only 46.37\% of the data is rejected and the average
bias reduces to -0.32\,K. When all the cloudy cases, except the ``too cloudy'',
are corrected, the total bias increases to 0.88\,K. The unusually high fraction
of cloudy cases and an increase in bias on correction suggests
misclassification of clear cases as cloudy. Since, the weighting functions of
AWS-32 and AWS-4X are poorly matched, their relationship is inadequate to
estimate the cloud impact in AWS-32. Without any additional information, the
regression approach can only provide a successful filtering in comparison to
the original approach described by \citet{rekha2012potential}, although at the
cost of high rejection fraction. Additional information seems necessary to improve
the accuracy of the correction.

\begin{figure}[!p]
	\centering
	\includegraphics[height=100mm]{PDF_corrected_AWS-32_AWS-4X_thinned}
	\caption{Same as Fig.~\ref{fig:correction:c34-42:thinned}, but for channel combinations AWS-32 and AWS-4X. The dataset ``corrected 2$\sigma$'' is not shown.  }
	\label{fig:correction:c32-4X:thinned}
\end{figure}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std & skewness & fraction \\
						&   [K]  &   [K] & [-] 		& rejected [\%]\\
		\hline
Noise             		&  0.00 &  0.45 &               0.01 &                - \\
uncorrected       		&  -2.25 &  8.94 &              -6.85 &                - \\
filtered($1\sigma$)  	&  -0.21 &  0.51 &              -0.65 &               85.43 \\
corrected($1\sigma$) 	&   0.88 &  1.76 &              -1.30 &                5.35 \\
corrected($2\sigma$) 	&   0.76 &  1.80 &              -1.07 &                5.35 \\
sreerekha et al   		&  -0.32 &  0.78 &              -1.94 &               46.37 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:correction:stats:34:42:thinned}, but for
      channel combinations: AWS-32 and AWS-4X. }
	\label{tab:correction:stats:32:4X}
\end{table}

\subsection{Application of correction scheme to AWS-32, AWS-33, AWS-35 and AWS-36}
%

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std & skewness & fraction \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		noise             		&   0.00 &  0.45 &               0.00 &                - \\
		uncorrected       		&  -2.25 &  8.94 &              -6.85 &                 - \\
		filtered($1\sigma$)  	&  -0.26 &  0.77 &             -16.33 &               72.06 \\
		corrected($1\sigma$) 	&   0.43 &  1.70 &             -14.68 &                7.93 \\
		corrected($2\sigma$) 	&   0.25 &  1.72 &             -13.86 &                7.93 \\
		 sreerekha et al   		&  -0.32 &  0.78 &              -1.94 &               46.37 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:correction:stats:34:42:thinned}, but for channel combination: AWS-32 and AWS-41.   }
	\label{tab:correction:stats:32:41}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &  skewness  & fraction\\
		&   [K]  &   [K] & [-]  & rejected [\%]\\
		\hline
		noise       	    	&  0.00 &  0.45 &               0.00 &               - \\
		uncorrected	    		&  -1.74 &  7.45 &              -7.77 &                - \\
		filtered($1\sigma$)  	&  -0.13 &  0.49 &              -0.40 &               26.18 \\
		corrected($1\sigma$) 	&  -0.05 &  0.59 &              -0.85 &                6.88 \\
		corrected($2\sigma$) 	&  -0.11 &  0.57 &              -1.17 &                6.88 \\
		sreerekha et al         &  -0.23 &  0.65 &              -1.52 &               46.37 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:correction:stats:34:42:thinned}, but for channel combination: AWS-33 and AWS-41.   }
	\label{tab:correction:stats:33:41}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise           	  &  0.00  &  0.63 &               0.00 &                -\\
		uncorrected     	  &  -0.83 &  4.50 &             -10.85 &                - \\
		filtered($1\sigma$)	  &  -0.13 &  0.68 &              -0.46 &               21.30 \\
		corrected($1\sigma$)  &   0.01 &  0.90 &              -1.65 &                2.86 \\
		corrected($2\sigma$)  &  -0.09 &  0.83 &              -2.35 &                2.86 \\
		sreerekha et al       &  -0.11 &  0.67 &              -0.22 &               46.37 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:correction:stats:34:42:thinned}, but for channel combination: AWS-35 and AWS-43.}
	\label{tab:correction:stats:35:43}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise            	 &  0.00 &  0.88 &              -0.01 &               - \\
		uncorrected      	 &  -0.54 &  3.40 &             -12.26 &                - \\
		filtered($1\sigma$)  &  -0.11 &  0.92 &              -0.13 &                7.38 \\
		corrected($1\sigma$) &  -0.07 &  0.93 &              -0.07 &                2.50 \\
		corrected($2\sigma$) &  -0.11 &  0.94 &              -0.19 &                2.50 \\
		sreerekha et al      &  -0.08 &  0.90 &              -0.02 &               46.37 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:correction:stats:34:42:thinned}, but for channel combination: AWS-36 and AWS-43.   }
	\label{tab:correction:stats:36:43}
\end{table}

The correction scheme can also be extended to other combinations of 183GHz and
325GHz channels. The most relevant combinations for clear-sky screening are: AWS-32 and
AWS-41; AWS-33 and AWS-41; AWS-35 and AWS-43; AWS-36 and AWS-43.

Table~\ref{tab:correction:stats:32:41} shows the statistics when AWS-32 is
corrected using AWS-41. The bias and standard deviation of the filtered dataset are better than
the approach by \cite{rekha2012potential}. However, the results from ``sreerekha et al'' have lower skewness. After correction, the corrected data have poorer
statistics. This is not surprising as the weighting function match between
AWS-32 and AWS-41 is poor. The correction using $2\sigma$ as threshold works
better $1\sigma$. This is an exception, the reverse is valid for AWS-34/42 (as
mentioned) and the channel combinations discussed below.

Table~\ref{tab:correction:stats:33:41} shows the statistics when AWS-33 is
corrected using AWS-41. The TB differences for this combination and the cloud
impact in AWS-33 are highly correlated. Rejecting the cloudy cases reduces the
average bias in the measurements from -1.74\,K to -0.13\,K, while a correction
of cloudy cases reduces the bias further to 0.05\,K. The standard deviation and
skewness of the corrected dataset are higher than pure-filtering, but smaller
than the approach by Sreerekha et al. Using $2\sigma$ threshold corrects a
smaller fraction of cases than $1\sigma$ but the overall statistics are poorer
in comparison to $1\sigma$, but still better than Sreerekha et al.

Table~\ref{tab:correction:stats:35:43} shows the statistics for the combination AWS-35 and AWS-43. Here the two filtered datasets have a comparable performance, while on correction, the bias is reduced, but skewness and standard deviation are increased. After correction ($1\sigma$) the bias due to cloudy measurements reduces from -0.83\,K to 0.01\,K while rejecting only 2.86\% of the total cases.  The performance of $1\sigma$ threshold in this case is also better than $2\sigma$.

Table~\ref{tab:correction:stats:36:43} shows the statistics when AWS-36 is
corrected using AWS-43. For this combination, the weighting functions have a
poor match, but the TB differences and the cloud impact are strongly
correlated, as the cloud impact in AWS-36 is relatively low in comparison to
other 183\,GHz channels. The low fraction of rejection (7.38\%) also shows that
most of the measurements are without cloud impact. However, the small fraction
of cloudy cases which are corrected, helps in successfully reducing the average
bias from -0.54\,K to -0.07\,K.

\section{Cloud correction by QRNN}
%
\label{sec:qrnn}
There are two main drawbacks of the simple scheme introduced in the previous
section. The filtering/correction only involves a single 325\,GHz channel. The
scheme could be extended to consider a higher number of channels, but there is
no obvious formulation of the expressions to apply. The second problem is that
no case-specific uncertainty estimate is provided, only the overall bias and
standard deviations are known. The first drawback hints towards using a machine
learning approach, and the second one suggests to use the special form denoted
QRNN (Quantile Regression Neural Network), that was introduced for atmospheric
retrievals by \citet{pfreundschuh:aneur:18}.

\subsection{QRNN model}
%
The neural network training is a process of learning to predict the outputs {$y_i$} from inputs {$x_i$} through a series of learnable transformations. While training, neural networks seek to minimise the model error through a loss function. The choice of the loss function depends on the predictive problem. When a neural network is trained to minimise the mean of the quantile loss function to predict the quantiles of the distribution, it is called a Quantile Regression Neural Network (QRNN).

If $x_{\tau}$ is the $\tau$th quantile of a cumulative distribution function $F(X)$, the quantile loss function is defined as:
\begin{equation}
\mathcal{L}_\tau(y_\tau, y_{true}) =
\begin{cases*}
(1 - \tau)|y_\tau - y_{true}| &\text{if}  $y_\tau < y_\text{true}$\\
\tau |y_\tau - y_\text{true}| & \text{ otherwise, }
\end{cases*}
\label{loss_function}
\end{equation}

A detailed description of QRNN can be found in \citet{pfreundschuh:aneur:18}.

\subsubsection{Training}
%  
\label{sec:qrnn:training}
\begin{figure}[!tb]
	\centering
	\includegraphics[height=85mm]{quantile_loss.png}
	\caption{The mean quantile loss with different configurations of layer width and hidden layers ($n_h$). Here QRNN is trained to predict clear-sky values for AWS-34 using four channels option of 325\,GHz}
	\label{fig:quantile_loss}
\end{figure}
In this section, QRNNs are used to correct the cloud impact in each of 183\,GHz
channels using TBs measurements from all available channels of 325\,GHz. A
separate QRNN is used for each of the 183\,GHz channels. For example, to
predict clear-sky TBs from AWS-34, the training data is measurements from all
channels of 325\,GHz and measurements from the 183\,GHz to be predicted. We
assess the performance of the scheme using both three channel and four channel
options of 325\,GHz. Two suggestions for three channel option are described in
Sec.~\ref{sec:channel:configurations}, and both have been used to train QRNN
separately. Thus, for each channel of 183\,GHz, we have a set of three
predictions: one from four channel option, and two from the old and new
configurations of the three channel option. Further in the text, the old and
new three channel configurations are labelled as "three-a" and "three-b".

To develop the QRNN scheme, we use the same dataset as in the
simple scheme . Out of 125\,000 simulations, 90\% are used to train the model,
while the rest 10\% is used to evaluate the model's performance during
training. Presently we train QRNN for only nadir sensor viewing angle and the
receiver noise $T_r$ = 1200\,K. In the training process, noise is added to the
input according to Eq.~\ref{eq:noise}, every time the training data are exposed
to the model. This ensures that the model is exposed to a new dataset on every
iteration. Before training, the input data is normalized with mean and standard
deviation.

The structure of the QRNN is defined by various hyper-parameters which are
tuned to achieve a high-performing model. These parameters determine the
structure of the neural network and how the network is trained. We performed a
grid search to determine structural parameters: depth (number of hidden layers)
and width (number of neurons). The quantile loss on the validation set averaged
over all predicted quantiles is used to compare the performance of different
hyper-parameter configurations. Figure~\ref{fig:quantile_loss} shows the mean
quantile loss when QRNN is trained for predicting clear-sky values for AWS-34
using four channel option. Increasing both layer width and depth has a positive
impact on performance of the network, however for four hidden layers,
increasing the layer width leads to only small performance gains. On basis of
these results, a neural network with four hidden layers and 128 neurons was
chosen. For the training, we obtained the best performance when trained with a
multiple learning rate schedule. We start the training process with a initial
learning rate and after a certain number of epochs, the training is updated
with a new initial learning rate. We chose three training schedules with
initial learning rates = 0.01, 0.001, 0.0001 respectively, updated after 20
epochs.

The grid search was also conducted for both three channel options, and an identical QRNN structure, as described for the four channel option, was found to have the best performance. For QRNN models presented here, we did not optimise the type of activation function and batch size. Rectified Linear Unit (ReLU) was used as the activation function  and the batch size was set to 256 samples for all QRNN models.

\subsection{Prediction accuracy}
%
We train the QRNN to predict the quantiles $x_{\tau}$, for median, $\pm 1\sigma$, $\pm 2 \sigma$ and  $\pm 3 \sigma$ ($\tau$ = 0.002, 0.03, 0.16, 0.5, 0.84, 0.97, 0.998). The expectation value of the posterior distribution is assumed to be the best estimate for the predicted clear-sky value. To analyse the prediction accuracy, a testing dataset is constructed by randomly drawing 25\,000 cases from the simulation dataset. Noise is added to each simulation to mimic a new measurement dataset.

The distributions of the predicted clear-sky noise-free values for each of the
183\, GHz channels are shown in Figure~\ref{fig:pdf:qrnn}. The simulated
clear-sky distributions are also plotted for reference. These results are from
the four channel option. The distributions show that QRNN is overall successful in
predicting the clear-sky cases. For all channels, the simulations and
predictions have a good agreement for the main peak of the distribution.
However, for AWS-32, AWS-33 and AWS-34, QRNN seems to predict slightly colder
TBs. These results give an indication of the behaviour of QRNN, but for a
detailed analysis of the prediction accuracy, we make a quantitative assessment
of the errors.

\begin{figure}[!p]
	\centering
	\includegraphics[height=60mm]{distribution_predicted_C32}\hspace{5mm}% 
	\includegraphics[height=60mm]{distribution_predicted_C33}
	\includegraphics[height=60mm]{distribution_predicted_C34}\hspace{5mm}% 
	\includegraphics[height=60mm]{distribution_predicted_C35}
	\includegraphics[height=60mm]{distribution_predicted_C36}
	\caption{ Probability distribution functions (PDFs) of predicted and observed noise free brightness temperatures for all 183\,GHz channels of AWS. The results are based on QRNN trained with four channel option of 325\,GHz.}
	\label{fig:pdf:qrnn}
\end{figure}
\begin{figure}[!p]
	\centering
	\includegraphics[height=85mm]{Channel_C34.png}
	\caption{Deviation of AWS-34 measurements to noise free clear-sky
		simulations before and after correction with QRNNs. Predictions from three channel (three-b) and four channel options for 325 GHz are shown. }
	\label{fig:qrnn_C34:deviations}
\end{figure}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise                       &  0.003 & 0.641 &             -0.001 &      - \\
		uncorrected                 & -0.677 & 3.673 &            -10.321 &      - \\
		corrected (three-a) 		&  0.094 & 0.624 &             -0.629 &      2.750 \\
		corrected (three-b) 		&  0.100 & 0.652 &             -0.543 &      2.582 \\
		corrected (four)   			&  0.111 & 0.620 &             -0.659 &      2.985 \\
		\hline
	\end{tabular}
	\caption{ Bias, standard deviation and measure of skewness for the error distributions of the dataset (AWS-34) shown in Figure~\ref{fig:qrnn_C34:deviations}. The labels "three-a", "three-b" correspond to the old and new suggestions of three channels options of 325\,GHz, and the label "four" represents the four channels option.}
	\label{tab:qrnn:C34}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise                       & -0.004 & 0.450 &             -0.030 &      - \\
		uncorrected                 & -1.343 & 5.908 &             -7.484 &      - \\
		corrected (three-a)  		&  0.019 & 0.889 &             -3.519 &      4.906 \\	 
		corrected (three-b)  		& -0.008 & 0.850 &             -3.326 &      4.926 \\			
		corrected (four)    		& -0.011 & 0.878 &             -3.030 &      4.930 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C34}, but for channel AWS-32.}
	\label{tab:qrnn:C32}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise                       & -0.005 & 0.450 &             -0.004 &      - \\
		uncorrected                 & -0.982 & 4.703 &             -8.778 &      - \\
		corrected (three-a) 		&  0.088 & 0.583 &             -0.959 &      4.197 \\
		corrected (three-b)  		& -0.030 & 0.612 &             -0.756 &      4.205 \\
		corrected (four)            &  0.106 & 0.591 &             -1.006 &      4.212 \\
		\hline

	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C34}, but for channel AWS-33.}
	\label{tab:qrnn:C33}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline

		Noise                      &  0.003 & 0.626 &             -0.009 &      - \\
		uncorrected                & -0.428 & 2.653 &            -12.492 &      - \\
		corrected (three-a) 	   &  0.060 & 0.610 &             -0.281 &      1.439 \\
		corrected (three-b) 	   &  0.097 & 0.618 &              0.011 &      1.331 \\
		corrected (four)   		   &  0.042 & 0.591 &             -0.351 &      1.821 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C34}, but for channel AWS-35.}
	\label{tab:qrnn:C35}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise                       &  0.000 & 0.888  &              0.005 &      - \\
		uncorrected                 & -0.269 & 2.059 &            -11.899 &      - \\
		corrected (three-a) 		&  0.062 & 0.797 &             -0.139 &      1.279 \\
		corrected (three-b) 		&  0.092 & 0.770 &             -0.075 &      1.124 \\
		corrected (four)   			&  0.056 & 0.779 &             -0.172 &      1.156 \\

		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C34}, but for channel AWS-36.}
	\label{tab:qrnn:C36}
\end{table}

In an analogy between the results from the simple correction scheme and QRNN, we perform a similar error distribution analysis as in Section~\ref{sec:simple}. Figure~\ref{fig:qrnn_C34:deviations} shows the error distributions of the measurements and predicted clear-sky TBs for AWS-34 using three (three-b) and four channel options of 325\,GHz. Measurements with cases with $Tb2 - Tb1 > -15$\,K are assumed to be too-cloudy and are filtered out. Here $Tb1$ denotes TBs from AWS-34 and and $Tb2$ is the paired counterpart from 325\,GHz, i.e., AWS-42.  Both the three channels and four channels QRNN models are successful in predicting the clear-sky noise-free value for AWS-34, and the resulting error distribution is symmetric. The statistics corresponding to these distributions are provided in Table~\ref{tab:qrnn:C34}. The fraction of cases rejected as "too-cloudy" is also provided under ``fraction rejected''. For AWS-34, average deviation in the predictions from all three QRNN models are comparable.  The cloud impact in the uncorrected measurements is reflected in the high negative skewness (-10.32), however the predictions have relatively symmetric error distributions as reflected by the low skewness values. Among the three set of predictions, option three-b has the most symmetric error distribution but with a relatively large spread. Predictions from option four have the lowest standard deviation (0.620\,K), which is interestingly smaller than the standard deviation of noise (0.641\,K). Though, all three options of 325\,GHz provide a full coverage to the hydrometer impact in AWS-34, predictions from four channel option have slightly better accuracy.

Similar error statistics are also estimated for other 183\,GHz channels.
Table~\ref{tab:qrnn:C32} provides the statistics for AWS-32. Before correction,
the average bias and standard deviation with respect to noise free simulations
is -1.343\,K and 5.908\,K respectively. However, after correction (four
channel), the bias and standard deviation are -0.011\,K and 0.878\,K. The
skewness in the error distributions is also reduced in the predictions (from
-7.48 to -3.03). But a relatively high negative skewness in the predicted
dataset indicates presence of uncorrected cloud impact. For AWS-32, option
``three-b'' has the lowest spread of errors, but the error distributions are
more symmetric for option ``four''.

Table~\ref{tab:qrnn:C33} shows the statistics for AWS-33. The standard
deviation in the four channel option (0.591\,K) is lower than
``three-b''(0.612\,K), but the skewness in three-b in lower. This indicates
that with four channel option, the QRNN predictions are mostly closer to the
simulations, but few cases with large cloud impact are under-predicted.

For AWS-35 (Table~\ref{tab:qrnn:C35}), the four channels option has the best performance. The average bias in QRNN predictions (four) is only 0.04\,K in comparison to -0.42\,K for the measurements. Option ``three-b'' has lower skewness than option four, but the standard deviation is higher. It is interesting to note that the standard deviation in all three predicted datasets is always lower than noise. Similar results are also seen for AWS-36 (Table~\ref{tab:qrnn:C36}). The four channels combination reduces the average bias from -0.269\,K to 0.056\,K and the standard deviation from 2.059\,K to 0.779\,K. Also for this channel, the spread of the predictions is narrower than noise.

\subsection{Uncertainty estimation}
%
\begin{figure}[!tb]
	\centering
	\includegraphics[height=80mm]{prediction_uncertainty.png}
	\caption{The prediction uncertainty in AWS-34 with respect to quantiles for randomly selected 2400 cases. The red line represents the uncertainty if underlying distribution is purely Gaussian with mean 270\,K and standard deviation 0.62\,K. }
	\label{fig:prediction_uncertainty}
\end{figure}

Quantile regression can be used to construct a probability distribution of the
predictions in contrast to simple regression models, which predict only a point
estimate of the target. In our QRNN model, we estimate the conditional
quantiles for median, $\pm 1\sigma$, $\pm 2\sigma$, and $\pm 3\sigma$. The
predictions for each quantile are used to construct the confidence intervals.
The confidence interval for each point estimate is a way to quantify its
uncertainty. For example, the predicted quantiles for $\pm 1 \sigma$ give 68\%
probability that the true value will lie between the range.

To inspect if the predictions follow a normal distribution we evaluate the
length of confidence intervals from the median.
Figure~\ref{fig:prediction_uncertainty} shows the prediction uncertainty in
AWS-34 (four channel option). The uncertainty with respect to the median for
1000 randomly selected cases is shown. The red line represents the uncertainty
in a hypothetical Gaussian distribution of $270\pm0.62$\,K. If the underlying
distribution is Gaussian, the prediction uncertainty should be symmetric around
the median, and the uncertainties should be equally spaced for $1\sigma$,
$2\sigma$ and so on. The predictions from our QRNN model seem to be symmetric
around the median for the 94\% confidence interval ($\pm 2 \sigma$). The figure
also shows that the uncertainties for all 1000 random cases are concentrated
along a narrow width, thus suggesting that the case to case variability of the
uncertainty is low. A similar uncertainty quantification for other 183\,GHz
channels also gives similar results.

\subsection{QRNN using AWS-4X}
%
\begin{figure}[!tb]
	\centering
	\includegraphics[height=80mm]{Channel_C32_C4X.png}
	\caption{ Error distributions for measured and predicted AWS-32, when QRNN  model is trained with measurements from AWS-4X and AWS-32. The results from QRNN model trained with four channels 325\,GHz are also shown for comparison. }
	\label{fig:qrnn_C32_C4X:deviations}
\end{figure}

\begin{table}[!tb]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
		Noise                      & -0.001 & 0.451 &              0.020 &      - \\
		uncorrected                & -1.343 & 5.908 &             -7.484 &      - \\
		corrected (AWS-4X)         &  0.008 & 1.514 &             -4.085 &      3.678 \\
		corrected (four channel)   & -0.011 & 0.878 &             -3.030 &      4.930 \\
		\hline
	\end{tabular}
	\caption{Statistics for the error distributions of AWS-32 shown in Figure~\ref{fig:qrnn_C32_C4X:deviations}. The QRNN model is trained with measurements from AWS-4X and AWS-32.}
	\label{tab:qrnn:C32:C4X}
\end{table} 
We also tested the QRNN model to predict cloud-free TBs by using measurements
from AWS-4X and AWS-32 as input. In this case, we trained QRNN with
measurements from AWS-32 and AWS-4X, to predict the clear-sky cloud free TBs
for AWS-32. The model was trained as described in
Section~\ref{sec:qrnn:training}, but slightly better training results were
obtained with width as 256 neurons. Other structural and training parameters
remain unchanged.

Figure~\ref{fig:qrnn_C32_C4X:deviations} shows the error distributions for the
differences between the predicted values and clear-sky simulations for AWS-32.
The corresponding statistics are shown in Table~\ref{tab:qrnn:C32:C4X}. The
results from QRNN model trained with four channel option of 325\,GHz are also
shown for comparison.

When trained with AWS-4X, the predictions are not very successful in
alleviating the cloud impact. Although, the error standard deviation of the
predicted dataset is lower, a large negative skewness (-4.085) indicates
significantly high uncorrected cloud impact. On the other hand, QRNN
predictions trained with 325\,GHz data are also not completely cloud free, but
have a slightly better accuracy. The error standard deviation for the four
channels option is 0.878\,K.
Clearly, without additional information, QRNN cannot learn to predict cloud
free values for AWS-32. As already seen in simple correction scheme, AWS-4X is
not sufficient enough to characterise the effect of hydrometeors in AWS-32.

\section{Sensitivity to different $T_r$}
\label{sec:tr}
%
In the previous section, we presented the results from QRNN trained using the
three suggestions of 325\,GHz proposed in
section~\ref{sec:channel:configurations}. The receiver noise ($T_r$) was
assumed to be 1200\,K for all cases. In this section, we explore the
sensitivity of QRNN to a higher receiver noise. The QRNN training is repeated
for all three suggestions of 325\,GHz, but with $T_r$ = 1800\,K and $T_r$ =
2400\,K. QRNN is trained with identical structural and training parameters as
described earlier. Tables showing the statistics for error distributions for
each channel are provided. Table~\ref{tab:qrnn:C32:Trec:1800} -
Table~\ref{tab:qrnn:C36:Trec:1800} show the statistics for AWS-32, AWS-33,
AWS-34, AWS-35 and AWS-36 respectively, when $T_r$ = 1800\,K. Similarly
Table~\ref{tab:qrnn:C32:Trec:2400} - Table~\ref{tab:qrnn:C36:Trec:2400} show
the statistics when $T_r$ = 2400\,K.

For all channels, increase in the measurement noise increases the standard
deviation of the predictions. For example, for AWS-34 (four channel option),
the error standard deviation is 0.620\,K, 0.630\,K and 0.643\,K for $T_r$ =
1200\,K, 1800\,K and 2400\,K respectively. It is difficult to gauge the impact
of increasing noise through bias and skewness of the error distributions, as
the distribution can have a large spread, but still be symmetric with low bias
and skewness. It is interesting to note that even for 1800\,K, the error
standard deviation for predicted values for channels AWS-34, AWS-35 and AWS-36
is smaller than noise. This is in agreement with the results seen for $T_r$ =
1200\,K. However, for $T_r$ = 2400\,K, only AWS-36 follows the trend.

Overall, the results show that an increase in the measurement noise does not have a huge impact on on QRNN accuracy. For the four channel option, with a 100\% increase in noise, the change in standard deviation is less than 10\%.

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
	Noise                      & -0.001 & 0.451 &              0.020 &      - \\
	uncorrected                & -1.342 & 5.908 &             -7.479 &      - \\
	corrected (three-a) 	   & -0.009 & 0.893 &             -3.092 &      4.990 \\
    corrected (three-b) 	   &  0.009 & 0.888 &             -3.197 &      4.902 \\	
	corrected (four channel)   &  0.006 & 0.860 &             -2.943 &      4.954 \\
		\hline
	\end{tabular}
	\caption{Error statistics for AWS-32, when the model is trained with $T_r$ = 1800\,K. The definitions of "three-a", "three-b" and "four" are same as described in Table~\ref{tab:qrnn:C34} }
	\label{tab:qrnn:C32:Trec:1800}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
	Noise                       & -0.001 & 0.452 &              0.015 &      - \\
	uncorrected                 & -0.975 & 4.703 &             -8.763 &      - \\
	corrected (three-a) 		&  0.085 & 0.587 &             -0.640 &      4.209 \\
	corrected (three-b) 		&  0.062 & 0.605 &             -0.845 &      4.205 \\
	corrected (four channel)    &  0.048 & 0.606 &             -1.112 &      4.264 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:1800}, but for AWS-33.}
	\label{tab:qrnn:C33:Trec:1800}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
	Noise                       &  0.001 & 0.632 &              0.021 &      - \\
	uncorrected                 & -0.676 & 3.686 &            -10.301 &      - \\
	corrected (three-a) 		&  0.079 & 0.646 &             -0.344 &      2.730 \\
    corrected (three-b) 		&  0.090 & 0.639 &             -0.405 &      2.582 \\
	corrected (four channel)    &  0.079 & 0.630 &             -0.399 &      2.981 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:qrnn:C32:Trec:1800}, but for AWS-34. }
	\label{tab:qrnn:C34:Trec:1800}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
Noise                       &  0.001 & 0.629 &              0.007 &      - \\
uncorrected                 & -0.417 & 2.654 &            -12.446 &      - \\
corrected (three-a)  		&  0.045 & 0.636 &             -0.188 &      1.431 \\
corrected (three-b)  		&  0.081 & 0.640 &             -0.458 &      1.299 \\
corrected (four channel)    &  0.058 & 0.617 &             -0.007 &      1.801 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:qrnn:C32:Trec:1800}, but for AWS-35. }
	\label{tab:qrnn:C35:Trec:1800}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
Noise                      &  0.002 & 0.882 &             -0.006 &      - \\
uncorrected                & -0.268 & 2.063 &            -12.003 &      - \\
corrected (three-a)        &  0.075 & 0.831 &             -0.130 &      1.283 \\
corrected (three-b)        &  0.105 & 0.807 &             -0.080 &      1.136 \\
corrected (four channel)   &  0.093 & 0.826 &             -0.175 &      1.120 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:1800}, but for AWS-36.}
	\label{tab:qrnn:C36:Trec:1800}
\end{table}

%
\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
Noise                       &  0.002 & 0.454 &              0.003 &      - \\
uncorrected                 & -1.343 & 5.907 &             -7.483 &      - \\
corrected (three-a)  		& -0.017 & 0.877 &             -3.381 &      4.934 \\
corrected (three-b)  		& -0.015 & 0.882 &             -3.177 &      4.974 \\
corrected (four channel)    &  0.028 & 0.876 &             -3.638 &      4.978 \\
		\hline
	\end{tabular}
	\caption{ Error statistics for AWS-32, when the model is trained with $T_r$ = 2400\,K. The definitions of "three-a", "three-b" and "four" are same as described in Table~\ref{tab:qrnn:C34}  }
	\label{tab:qrnn:C32:Trec:2400}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
 Noise                     & -0.003 & 0.448 &              0.005 &      - \\
uncorrected                & -0.979 & 4.709 &             -8.762 &      - \\
corrected (three-a) &  0.057 & 0.628 &             -0.728 &      4.232 \\
corrected (three-b) &  0.040 & 0.614 &             -0.786 &      4.205 \\
corrected (four channel)   &  0.043 & 0.614 &             -0.798 &      4.232 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:2400}, but for AWS-33.  }
	\label{tab:qrnn:C33:Trec:2400}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
 Noise                      & -0.001 & 0.638 &              0.006 &      - \\
uncorrected                 & -0.676 & 3.688 &            -10.299 &      - \\
 corrected (three-a) &  0.069 & 0.655 &             -0.396 &      2.666 \\
corrected (three-b)  &  0.090 & 0.648 &             -0.310 &      2.594 \\
corrected (four channel)    &  0.059 & 0.643 &             -0.340 &      2.937 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:2400}, but for AWS-34.  }
	\label{tab:qrnn:C34:Trec:2400}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
Noise                      & -0.004 & 0.633 &              0.021 &      - \\
uncorrected                & -0.420 & 2.654 &            -12.449 &      - \\
corrected (three-a) 	   &  0.034 & 0.646 &             -0.389 &      1.479 \\
corrected (three-b) 	   &  0.076 & 0.647 &             -0.148 &      1.291 \\
corrected (four channel)   &  0.044 & 0.632 &             -0.188 &      1.809 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:2400}, but for AWS-35. }
	\label{tab:qrnn:C35:Trec:2400}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
		Dataset  		  &   bias &   std &   skewness & fraction  \\
		&   [K]  &   [K] & [-] & rejected [\%]\\
		\hline
Noise                      &  0.005 & 0.884 &             -0.026 &      - \\
uncorrected                & -0.268 & 2.057 &            -11.977 &      - \\
corrected (three-a) &  0.039 & 0.854 &             -0.131 &      1.283 \\
corrected (three-b) &  0.090 & 0.839 &             -0.096 &      1.108 \\
corrected (four channel)   &  0.059 & 0.850 &             -0.140 &      1.140 \\
		\hline
	\end{tabular}
	\caption{ Same as Table~\ref{tab:qrnn:C32:Trec:2400}, but for AWS-36.  }
	\label{tab:qrnn:C36:Trec:2400}
\end{table}

\section{Degrees of freedom}
%
An information content analysis is made by estimating the degrees of freedom
(DoF) of the simulated measurements. The analysis is similar to the one
described by \citet{eriksson:towar:20} for ICI.

Let $\mathbf{Y}$ be an ensemble of noise-free measurements, and $\textbf{E}$
and $\mathbf{\Lambda}$ be the corresponding eigenvectors and eigenvalues; then
in the eigenvalue space, the covariance matrix $\mathbf{S_y}$ can be written as

\begin{equation}
\mathbf{S_y = E \Lambda E^T} 
\end{equation}
Similarly, the uncertainty due to thermal noise is
\begin{equation}
\mathbf{S_{\Lambda} = E S_{\epsilon} E^T }
\end{equation}
where, $\mathbf{S_{\epsilon}}$ is a diagonal matrix with NE$\Delta$T$^2$ as
diagonal elements. The DoF is the number of diagonal elements of
$\mathbf{\Lambda}$ greater than the $\mathbf{S_\Lambda}$. For further details
see \citet{eriksson:towar:20}.

To calculate the DoF, the ensemble of simulations from 89\,GHz, 166\,GHz,
183\,GHz are used along with 229\,GHz or 325\,GHz. In order to compare the DoF
for cloudy and clear cases, we divide the database into three parts: ``all'',
``clear'' and ``cloudy'. All cases with cloud impact greater than 1\,K in any
of the 183\,GHz channels are classified as ``cloudy'', otherwise they are
classified as ``clear''. With this criteria, out of total 125460 cases, 12905
are classified as cloudy.

As example, Fig~\ref{fig:DoF:clear:1200K} shows the ratio of
$\Lambda/S_\Lambda$ with respect to the eigenvalue index for clear-sky cases.
The DoF below is an interpolation between the last index having a ratio above
one and next index.

\begin{figure}[!tb]
	\centering
	\includegraphics[height=80mm]{DoF.png}
	\caption{ The ratio of $\Lambda/S_\Lambda$ with respect to the eigenvalue
      index for clear-sky cases. The black dotted line represents the level
      where the ratio is unity. }
	\label{fig:DoF:clear:1200K}
\end{figure}

\subsection{DoFs for nominal IF range}

At this point we got feedback from Omnisys that also channel option 3b is
technically challenging. In short, it is significantly more difficult to
implement a 3\,GHz wide channel at an IF of 3.3\,GHz (AWS-42) than compared to
at 6.5\,GHz (AWS-41). To decrease the ratio between channel widths and centre
IF, a third three channel option was formulated (Table~\ref{tab:chs:3c}) that
is below denoted as three c. This option uses the same IF range as the
previously suggested ones.

\begin{table}[!p]
  \centering  
  \begin{tabular}[c]{c|c|c|ccc}
    Channel & Frequency   & Bandwidth & $T_r=1200$\,K
    & $T_r=1800$\,K & $T_r=2400$\,K\\
    name    & [GHz] &  [MHz] & [K]  & [K] & [K]\\
    \hline
    AWS-41  & 325.15$\pm$6.25 & 3500  & 0.54 & 0.76 & 0.98\\
    AWS-42  & 325.15$\pm$3.05 & 2500  & 0.64 & 0.90 & 1.16\\
    AWS-43  & 325.15$\pm$1.20 & \phantom{0}800 & 1.12 & 1.59 & 2.06\\
    \hline
  \end{tabular}
  \caption{The third three channel suggestion (three-c).}
  \label{tab:chs:3c}
\end{table}

Table~\ref{tab:DoF:all} shows the DoF with 325\,GHz when all cases are
considered. For $T_r$ = 1200\,K, the DoF from all four options is around 8,
however, increase in receiver noise leads to a loss in the DoF.
Table~\ref{tab:DoF:clear} and Table~\ref{tab:DoF:cloudy} show the DoF for
clear-sky and cloudy cases respectively. For clear-sky situations, the DoFs are
about 77. With increase in $T_r$, the DoF are degraded. For $T_r$ = 2400\,K, we
have only 5 DoF for option three-a and 4.8 for both option three-b/c and 4.9
for option four. On the other hand, the presence of clouds increases the DoF
significantly. For all three options, we have DoF close to 9. Increase in noise
has a lower effect on the DoF for cloudy cases.

The DoF values should not be interpreted to strongly. There appears to be some
``noise'' in the DoF, especially for higher noise. For example, in
Table~\ref{tab:DoF:all} all the options have fairly similar DoF for
$T_r=1800$\,K while for $T_r=2400$\,K there are considerable differences. This
shift when going $T_r=2400$\,K is not observed in the two other tables. Having
this in mind, it seems that options three b and c give very similar DoF. If
something, option three-c seems slightly better. The four channel option
obtains somewhat lower DoF than the three channel ones.

Table~\ref{tab:DoF:4X} gives the DoF, with 229\,GHz channel. The DoF is 6.8 for
both clear and cloudy situations. This is consistent with the fact that
229\,GHz has no redundancy with the five 183\,GHz channels, but there exists
some redundancy with the two window channels (89 and 166\,GHz).

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c}
		 325\,GHz & \multicolumn{3}{c}{DoF}   \\
		 option &  $T_r = 1200 $\,K &   $T_r = 1800 $\,K &   $T_r = 2400$\,K \\
		\hline
		three-a &              8.3 &              7.9 &              7.8 \\		
		three-b &              8.5 &              8.2 &              6.0 \\		
		three-c &              8.5 &              8.2 &              7.7 \\		
		four    &              8.2 &              7.8 &              6.2 \\
		\hline
	\end{tabular}
	\caption{The degrees of freedom for AWS with 325\,GHz channel. Both clear
      and cloudy cases are considered. Results for all three options of
      325\,GHz and different $T_r$ are shown. }
	\label{tab:DoF:all}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c}
	325\,GHz & \multicolumn{3}{c}{DoF}   \\
	option &  $T_r = 1200 $\,K &   $T_r = 1800 $\,K &   $T_r = 2400$\,K \\
	\hline
		three-a &              7.3 &              7.1 &              5.0 \\		
		three-b &              7.0 &              6.7 &              4.8 \\		
		three-c &              7.0 &              6.6 &              4.8 \\		
		four    &              7.0 &              6.0 &              4.9 \\
	\hline
\end{tabular}
	\caption{Same as Table~\ref{tab:DoF:all}, but only for clear cases.}
\label{tab:DoF:clear}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c|c|c}
	325\,GHz & \multicolumn{3}{c}{DoF}   \\
	option &  $T_r = 1200 $\,K &   $T_r = 1800 $\,K &   $T_r = 2400$\,K \\
		\hline
		three-a &              8.9 &              8.7 &              8.5 \\		
		three-b &              8.9 &              8.7 &              8.6 \\		
		three-c &              8.9 &              8.8 &              8.7 \\		
		four    &              8.7 &              8.5 &              8.3 \\
		\hline
	\end{tabular}
	\caption{Same as Table~\ref{tab:DoF:all}, but only for cloudy cases.  }
	\label{tab:DoF:cloudy}
\end{table}

\begin{table}[!p]
	\centering
	\begin{tabular}[b]{c|c}
		&   DoF \\
		\hline
		all    &      6.8        \\
		clear  &      6.8        \\
		cloudy &      6.8        \\
		\hline
	\end{tabular}
	\caption{The degrees of freedom for AWS with 229\,GHz channel. }
	\label{tab:DoF:4X}
\end{table}

\subsection{Impact of extending the IF range}
%
As a basic investigation of the impact of changing the width of the IF range, a
version of the four channel option was considered, denoted as four-b. It was
modified to cover 0.7 to 9.0 GHz (instead of 0.7 to 8.0 GHz). To focus on the
extension, only the edge channels were changed. That is, the revised AWS-41 and
have bandwidth of 3.8 and 0.9\,GHz, respectively. The DoFs obtained for four-b
are found in Table~\ref{tab:DoF:4b}.

As expected a slight improvement of the DoF can be seen for four-b. The
exception ``all'' and $T_r=2400$\,K, again likely an indication on some noise
in the DoFs coming out of the analysis.

That is, a broader IF range should give slightly better performance. If a
broader IF range is feasible, the position and width of all channels should be
revised, bit this is out of the scope of this study and anyhow hardly motivated
before it is known more exactly to what extent the IF range can be extended.

\begin{table}[!bt]
	\centering
	\begin{tabular}[b]{c|c|c|c|c}
      &	325\,GHz & \multicolumn{3}{c}{DoF}   \\
      &	option &  $T_r = 1200 $\,K &   $T_r = 1800 $\,K &   $T_r = 2400$\,K \\
      \hline
      all &	four   &              8.2 &              7.8 &              6.2 \\
      & four-a   &              8.3 &              8.0   &              6.1 \\

      \hline
      clear  &	four  &                7.0 &              6.0 &              4.9 \\
      &   four-a &                7.0 &              6.4 &              4.9 \\

      \hline	
      cloudy &	four  &              8.7 &              8.5 &              8.3 \\
      &   four-a &              8.8 &              8.6 &              8.4 \\
      \hline
\end{tabular}
\caption{The degrees of freedom for channel option four-b. Values for option
  four are repeated from Tables~\ref{tab:DoF:all}-\ref{tab:DoF:clear}. }
\label{tab:DoF:4b}
\end{table}
\section{Discussion}

\subsection{229 vs.\ 325\,GHz}
\subsubsection{Altitude coverage}
%
This discussion is based on figures and text found in Sec.~\ref{sec:overview}.
A basic remark is that a channel at 229\,GHz channel has a lower sounding
altitude than all the 183\,GHz channels for clear-sky conditions. Accordingly,
there are no matching RH weighting functions between 229\,GHz and 183\,GHz
channels. On the other hand, channels around 325\,GHz have quite similar
characteristics as some of the 183\,GHz channels (for clear-sky), but
consistently have a sounding altitude above the one of AWS-32.

With respect to sensitivity to hydrometeors, 229\,GHz will react on rain and
clouds at lower altitudes than the 183\,GHz channels. Due to the lower sounding
altitude, 229\,GHz is also more influenced by the surface. Both these aspects
could be beneficial in a wider perspective, but are not ideal for cloud
filtering of 183\,GHz data. With respect to this application, the main drawback
of the 325\,GHz region is the lack of a full coverage for AWS-32. There exists
an altitude range where hydrometeors have an impact on AWS-32 but will largely
be undetected by the 325\,GHz channels.

\subsubsection{Performance of the basic scheme}
%
A simple approach for cloud filtering/correction based on 325\,GHz data is
tested in Sec.~\ref{sec:simple}. The approach requires simulations for deriving
some parameters (in the form of a polynomial fit), while the application only
involves measurement data. Another advantage of the approach is that it is
channel specific. That is, it could result in that e.g.\ AWS-32 and 33 are
rejected, while that data for the higher peaking channels still can be used.

The results for AWS-33, 34, 35 and 36 should be considered as satisfactorily.
In short, a relatively low rejection rate is achieved and still data with
relatively narrow and symmetric error distributions are provided. The lowest
standard deviation and skewness are obtained by just performing a filtering.
Activating the correction part (with $1\sigma$ as threshold value) decreases
the bias and rejection rates significantly, but comes with the cost of
increased standard deviation and skewness. However, the increase in standard
deviation is throughout below 50\% (compared to the ideal case of that the
error distribution only consists of noise).

For AWS-32, filtering based either AWS-4X or AWS-41 has a very high rejection
rate, though the resulting dataset has low error. The high fraction of
rejection can increase the probability of removing ``true'' clear cases. After
correction, the error distributions are asymmetric and have large bias and
standard deviation. Among the two combinations, the dataset corrected with
AWS-41 has slightly better bias and standard deviation. As an exception to
other 183\,GHz channels, the correction scheme for AWS-32 works better for
$2\sigma$ threshold, as with a larger value of noise, relatively smaller number
of cases are corrected. The effect of mismatch between weighting functions for
AWS-32 and AWS-4X/41 is reflected in the performance of our correction scheme.

\subsubsection{Comparison to \citet{rekha2012potential}}
%
Results from \citet{rekha2012potential} has been used as reference, but the
comparison should only be considered as being qualitative. A more detailed
comparison is not possible as the selection of atmospheric cases is very
different between the studies. While in this study we aim to have a statistical
distribution as close as possible to real conditions, there is a strong focus on
cloudy conditions in the simulations used by \citet{rekha2012potential}. Our
translation of the performance reported in \citet{rekha2012potential} to our
conditions is found in Sec.~\ref{sec:rekha}.

The results of \citet{rekha2012potential} are better than ours for AWS-32, while
for the other 183\,GHz channels better results are obtained by using 325\,GHz
data. However, when only pure-filtering is considered, our scheme performs
better for AWS-32, albeit with a high rejection rate. For all later channels of
183\,GHz, we reject only a small fraction of cases in comparison to
\citet{rekha2012potential}, but the corrected datasets have a more symmetric error
distributions with lower bias and standard deviation. Only for AWS-35 and AWS-36,
the standard deviation and skewness of the corrected dataset are higher.

It should be noted that the approach of \citet{rekha2012potential} involves
simulations for every case, to obtain a ``background minus observation''
difference, while in this study a similar or better performance is obtained
from using measurement data alone.

\subsubsection{Performance of QRNN}
In section~\ref{sec:qrnn} we use QRNN as a method for cloud correction based on
325\,GHz data. Measurements from all 325\,GHz channels and the 183\,GHz channel
to be predicted are used to train the QRNN model.

The results demonstrate that QRNN is successful in predicting the clear-sky
noise-free data for AWS-33, AWS-34, AWS-35 and AWS-36. The error distributions
for the predictions are symmetric and have low bias and standard deviation. In
fact, for AWS-34, AWS-35 and AWS-36, the spread of errors is even smaller than
the noise. This is the case as a low fraction of the measurements have a
significant cloud impact, and for clear sky situations the QRNN prediction can
largely be seen as a weighted mean between the two channels. This averaging
cancels out some of the noise in the two channels. QRNN seems to perform
equally well with three configurations suggested for 325\,GHz channels.

For AWS-32, the QRNN trained with 325\,GHz is only partially successful in
reducing the impact of hydrometeors. The resulting error distributions have
significantly lower spread than the measurements, but are still negatively
skewed. Among the three combinations of 325\,GHz, the four channel option has
slightly better performance. Since 325\,GHz cannot provide a full coverage for
hydrometeor impact in AWS-32, QRNN cannot learn to predict cloud free values
for instances which lie outside the area of coverage. In order to compare the
229\,GHz channel with 325\,GHz, QRNN was also trained with AWS-4X to predict
cloud free AWS-32 TBs. The performance of QRNN trained with AWS-4X is worse
than training with 325\,GHz. The spread of error distributions is 1.5 times
more than obtained with four channels option of 325\,GHz.

\subsection{Technical recommendations}
%
The cloud correction and filtering results gave no clear indications on the
requirements of the channels, and as a complement also the measurements' degree
of freedom (DoF) were assessed. The later analysis gave a bit clearer
indications and the discussion in this section is mainly based on the DoF results.

\subsubsection{NE$\Delta$T requirement}
%
No critical dependency of the cloud correction to the 325\,GHz receiver noise
temperature ($T_r$) was found (Sec.~\ref{sec:tr}). For example, for AWS-36 and
the four channel option, the standard deviation only increased 0.85 from 0.7
when increasing $T_r$ to 2400\,K from 1200\,K. The DoF showed a higher
sensitivity to $T_r$, especially when going from 1800\,K to 2400\,K and only
considering clear cases (Table~\ref{tab:DoF:clear}).

ICI-5 has a width of 3\,GHz and has NE$\Delta$T requirement of 1.2\,K
\citep{eriksson:towar:20}. From e.g.\ Table~\ref{tab:chs:3b} it can be
understood that this scales to a requirement on $T_r$ of about 2700\,K (as
3\,GHz bandwidth and $T_r=2400$\,K gives a noise of 1.06\,K).

Combining the observations drawn in the two last paragraphs it seems reasonable
to set the threshold, target and breakthrough for NE$\Delta$T of AWS 325\,GHz
channels to match a $T_r$ of 2400, 1800 and 1200\,K, respectively. The
threshold value provide a sufficient basis for an acceptable cloud correction
and still represents a step forward compared to ICI. However, if the present
ambition of Omnisys of $T_r=1200$\,K will be reached, AWS will provide much
less noisy 325\,GHz observations than ICI, representing a breakthrough.

\subsubsection{Channel specifications}
%
We formed three first 325\,GHz channel options. All three options were
considered only in the QRNN part as it is difficult to maintain the full
information of all channels in a simple regression based scheme. None of the
three options gave consistently best results, and it is not possible to give a
clear recommendation. For this reason we did not see any value in testing
further options in this part.

Some further channel options were included on the DoF analysis. Staying inside
0.8\,-\,8.0\,GHz in terms of IF, the option three-c was found to have the
highest overall DoF, but the differences to three-b and four are relatively
small. However, it could be challenging to implement any of the three channel
options. Option three-a risks to be affected by interference of the downlink
and in options three-b and c there is a high ratio between bandwidth and centre
IF of AWS-42. This aspect also makes it hard to extend options three-b and c if
it would turn out that the IF range can be extended. A brief analysis showed
that a broader IF range gives some increase in DoF, that potentially could
improve the merit of the four channel option.

We don't see how these results can be boiled down to threshold, target and
breakthrough requirements. Option 3c came out best here, but the full
optimisation of the options was made and a modified or extended four channel
version could reach the same performance. In any case, the differences between
the options are small and it seems reasonable to allow that  the technically
most simple option is selected (that in fact can be the four channel one).
When this selection is done and the technical constraints are better known (IF
range available, needed frequent gaps between channels), an iteration of DoF
analysis should be made to see if some ``optimisation'' of the channel
specifications can be achieved.

Further, the simulations have assumed rectangular shaped response functions.
This was done for simplicity and shall in no way be taken as an indication on
that this shape is required. We have left some frequency gaps between the
channels to simplify the technical implementation, i.e.\ to decrease the demand
on very sharp transitions from high to low response.

\subsubsection{Polarisation response}
%
Polarisation aspects have not been considered in full detail. Some polarisation
is induced by the surface in the simulations, but hydrometeors are assumed to
be totally randomly oriented and do not cause polarisation. To involve oriented
particles was not possible due to the limited time allocated for the study.

The basic recommendation would be to use the same polarisation response as far
as possible. If both quasi-horizontal and quasi-vertical bands must be applied,
the recommendation prioritise to use the polarisation option for 183 and
325\,GHz, to maintain a similar influence of polarisation between the bands.

\section{Conclusion}
%
This study compares the options of having a channel at 229\,GHz or having some
around 325\,GHz from a single perspective, cloud filtering/correction of
183\,GHz data. There could be other important reasons to consider for the
selection.

It is demonstrated that the 325\,GHz channels are capable of providing a more
accurate cloud correction for 183\,GHz than the 229 \,GHz channel. The simple
regression based approach evaluated is channel specific and involves the
325\,GHz channel one at the time. In the second correction approach based on
machine learning, QRNN is used to predict the clear-sky noise-free values
throughout using all 325\,GHz in parallel. This seems to be a very useful
approach. The QRNN predictions result in error distributions with low skewness
and widths. In some case the prediction uncertainty is even lower than the
noise of the 183\,GHz channel targeted. This approach should be ideal for
making use of AWS in assimilation systems of clear-sky type.

No results gave strong constraints for defining threshold, target and
breakthrough requirements, but a suggestion is still given with respect to
receiver noise temperature. The three requirements levels are suggested to be
placed at 2400, 1800 and 1200\,K. On the other hand, the differences found
between the tested three and four channel options are too small to motivate any
clear recommendation. It seems reasonable to instead select between having
three or four channels based on the technical challenges associated with each
option and later setting the exact channel specifications.

{\footnotesize
\bibliography{j_abbr,references}
}

\end{document}